---
title: "Class06.R"
author: "Jiselle Arevalo"
format: pdf
---

> *Q1: Write a function add() to add some numbers together. Make sure to add comment(s) to the function explaining why you do things.*

```{r}
# This is subpart a. used to show the add() function (basic addition),
# adding 4 and 7 together

add_2<- function (x,y) {x + y}
add_2(4,7)

# This is subopart b. Is used to demonstrate basic addition of
# various numbers or variables, using sum because adding multiple numbers
add_multi <- function(x,...){ sum(x,...)}
add_multi(1,2,3,-4)
```

> *Q2:Write a function generate_dna() that generates a random DNA sequence of a length input by the user; e.g. generate_dna(11) may return "ATTTAAGGCTG". Make sure to add comments to the function. Play with functions sample() and paste() in your R-Studio Console to see how they can help here.*

```{r}
generate_dna <- function(x) { # generating a nucleotide vector
  nucs<- c("A","G","T","C")
  # using sample to pull out random nucleotides
  nucs_vector<- sample(nucs, x, replace=TRUE)
  # Generating a dna string from nucs_vector
  paste(nucs_vector, collapse="")
}
# this is testing my generate_dna() function
generate_dna(34)
```

> *Q3:Write a function generate_protein() that generates a random peptide/protein sequence of a length input by the user; e.g. generate_protein(6) may return "WQRTAG". Make sure to use the single-letter code for all 20 amino acids and use no other letters (see list of amino acids at the bottom of the page); make sure that individual amino acids can appear more than once. Make sure to add comments to the function.*

```{r}
generate_protein <- function(x) { # generating an amino acid vector
  amino <- c("A", "R", "N", "D", "C", "E", "Q", "G", "H", "I", "L", "K",
             "M", "F", "P", "S", "T", "W", "Y", "V")
  # using sample to pull out random amino acids
  
  amino_vector<- sample(amino, x, replace=TRUE)
                        
  # Generating a dna string from amino_vector
  
  paste(amino_vector, collapse="")
}
# testing my generate_protein() function!!
generate_protein(34)
```

> *Q4: Use your generate_protein() function to generate peptides of lengths 5, 7, 9, and 11 amino acids. Take advantage of the base R function sapply() for this*

```{r}
# making a vector for the sequence lengths to add into the sapply.
h <-c(5,7,9,11)
sapply(h,generate_protein, simplify = TRUE)
```

> *Q5: Take each of your peptides from Q4 and run a BLASTP search (https://blast.ncbi.nlm.nih.gov/Blast.cgi) selecting the Non-redundant protein sequences (nr) database and homo sapiens (taxid: 9606) organism. For each search, before clicking BLAST check the Show results in a new window tab next to the BLAST button to easily allow you to run the blast search for all four peptides at the same time in four different tabs. Once each search is done, for the top aligned protein for each peptide, list the percent identity (Per. Ident) times coverage (Query Cover)...*

```{r}
# this is for sequence 1 (YSRYF)


# This is query cover
100/100
# This percentage identity
100/100

#This is per ident x query
1*1*100

#100% is the answer for the sequence with 5 AA in pep.
```

```{r}
#intials are JAA

# this is for sequence 2 (WRLYWVP)


# This is query cover
71/100
# This percentage identity
100/100

#This is per ident x query
0.71*1*100

#71% is the answer for the sequence with 7 AA in pep
```

```{r}
# this is for sequence 3 (LIATKIIWK)


# This is query cover
67/100
# This percentage identity
100/100

#This is per ident x query
0.67*1*100

#67% is the answer for the sequence with 9 AA in pep
```

```{r}
# this is for sequence 4 (LECMTDELNMR)


# This is query cover
64/100
# This percentage identity
85.71/100

#This is per ident x query
0.64*0.8571*100

#54.8544% is the answer for the sequence with 9 AA in pep
```

> *Q6: Once there is a good amount of data in the shared excel sheet, export the sheet as a CSV UTF-8 file (Export option under the File tab) and save the file to your class working folder (i.e. the folder in which you are currently generating a Quarto document). In R-Studio, upload the csv file into a data frame using the command peptide_df \<- read.csv("your_filename"). Use the head() function to list the first 10 rows (note that the default for head() is 6) of the peptide_df dataframe as a test that everything looks good.*

```{r}
#Bimm peptide sheet was transfered into Rstudio files to be used for  the peptide data frame
peptide_df <- read.csv("BIMM143_peptide(Sheet1).csv")

#head() used to test table looks correct
head(peptide_df,10)
```

> *Q7:Use ggplot2 to generate a box plot with Length_aa on the x-axis and Max_Identity_perc on the y-axis.*

```{r}
library(ggplot2)
library(dplyr)
#assigning axis to certain data
x<-as.factor(peptide_df$Length_aa)
y<-(peptide_df$Max_Identity_perc)

#creating data plot, getting rid of outliers and adding labels to graph to make
#it more cleaner

ggplot(peptide_df) + aes(x,y)+ geom_boxplot (outliers=FALSE) + 
  labs(x="Length_aa",y="Max_Identity_perc")

#Assigning ggplot code to 'p' to make it more easier to access 
p<- ggplot(peptide_df) + aes(x,y)+ geom_boxplot (outliers=FALSE) + 
  labs(x="Length_aa",y="Max_Identity_perc")
```

> *Q8: Let's add a geom_jitter() layer to your box plot to visualize the individual data points on top of the box plot and highlight your own data points.*

```{r}
#adding geom_jitter layer to boxplot, this is subpart a!
p+geom_jitter(aes(col=Your_Initials))

```

```{r}
#this is subpart b for question 8 part i.
color <- c(rep("GREY",172))
table(color)
```

```{r}
#this is subpart b for question 8 part ii.
mydata<- peptide_df$Your_Initials=="JAA" # creating another vector for simplicity
table(mydata)
```

```{r}
# this is subpart b for question 8 part iii
color[mydata]<- "ROYALBLUE1" #this is for my peptides to stand out
table(color)
```

```{r}
# this is for question 8 subpart b, part iv.
p+geom_jitter(col=color)
```

> *Q8. PART C: Based on your boxplot, does any of your own peptide data fall outside the 25 to 75 percentile of the class's data? If yes, list which one(s).*

> **Answer: For the sequence length of 7, my peptide data falls outside of the 25-75 percentile for the class data\* \# its very hard to tell for length 7 as it is behind another point but if you look closely you can see it peek out!**

> *Q9:MHC class II molecules of our immune system display 9 amino acid peptides on the surface of immune cells that, when from a foreign origin, activates a T-cell immune response. Based on your findings in Q7 and 8, speculate why those peptides are 9 amino acids long rather than, say, 5 amino acids.*

> **Answer:Peptides that fall under the 9 AA are aren't very distinguishable and therefore will not be trigger an immune response. They are a perfect match and will not be seen under immune cell machinery. We see this on the boxplot as there is an 100% identity. We see that there is a large degree of variability through the large boxplot for 9aa. Meaning that 9 aa has more variety and therefore far more distinguishable, allowing immune response machinery to differentiate foreign matter from self matter. **
